\section*{Chapter 6}
\addcontentsline{toc}{section}{Chapter 6}

\subsection*{6.15 Show the result of each of the sample queries in Section 6.5 as it would apply to the database state in Figure 3.6.} 
\addcontentsline{toc}{subsection}{6.15}
\subsubsection*{Query 1: Find the name and address of all employees who work for the 'Research' department.}
\begin{center}
\begin{tabular}{ c | c | c }
  FNAME & LNAME & ADDRESS \\ \hline
  John & Smith & 731 Fondren, Houston, TX \\
  Franklin & Wong & 638 Voss, Houston, TX \\
  Ramesh &  Narayan & 975 Fire Oak, Humble, TX \\
  Joyce & English & 5631 Rice, Houston, TX \\
\end{tabular}
\end{center}

\subsubsection*{Query 2: For every project located in 'Stafford', list the project number, the controlling department number, and the department manager's last name, address, and birth date.}
\begin{center} 
\begin{tabular}{ c | c | c | c | c}
  PNUMBER & DNUM & LNAME & ADDRESS & BDATE \\ \hline
  10 & 4 & Wallace & 291 Berry, Bellaire, TX &20-JUN-31 \\
  30 & 4 & Wallace & 291 Berry, Bellaire, TX & 20-JUN-31 \\ 
\end{tabular}
\end{center}

\subsubsection*{Query 3: Find the names of all employees who work on all the projects controlled by department number 5.}
\begin{center}
\begin{tabular}{ c | c }
  LNAME & FNAME \\ \hline
\end{tabular}
\end{center}

\subsubsection*{Query 4: Make a list of project numbers for projects that involve an employee whose last name is 'Smith' as a worker or as a manager of the department that controls the project.}
\begin{center}
\begin{tabular}{ c }
  PNO \\ \hline
  1 \\
  2 \\
\end{tabular}
\end{center}

\subsubsection*{Query 5: List the names of all employees with two or more dependents.}
\begin{center}
\begin{tabular}{ c | c }
  LNAME & FNAME \\ \hline
  Smith & John \\
  Wong & Franklin \\
\end{tabular}
\end{center}

\subsubsection*{Query 6: List the names of employees who have no dependents.}
\begin{center}
\begin{tabular}{ c | c }
  LNAME & FNAME \\ \hline
  Zelaya & Alicia \\
  Narayan & Ramesh \\
  English & Joyce \\
  Jabbar & Ahmad \\
  Borg & James \\
\end{tabular}
\end{center}

\subsubsection*{Query 7: List the names of managers who have at least one dependent.}
\begin{center}
\begin{tabular}{ c | c }
  LNAME & FNAME \\ \hline
  Wallace & Jennifer \\
  Wong & Franklin \\
\end{tabular}
\end{center}

\subsection*{6.16 Specify the following queries on the COMPANY relational database schema shown in Figure 5.5, using the relational operators discussed in this chapter. Also show the result of each query as it would apply to the database state in Figure 3.6.}
\addcontentsline{toc}{subsection}{6.16}
I use the symbol $\sigma$ for SELECT, $\pi$ for PROJECT, $\rhd\lhd$ for EQUIJOIN, * for NATURAL JOIN, and $\mathfrak{F}$ for FUNCTION.

\subsubsection*{(a) Retrieve the names of employees in department 5 who work more than 10 hours per week on the "ProductX" project.}
\subsubsection*{Relational Operators}
EMP\_W\_X $\Leftarrow$ ( $\sigma$ \textsubscript{PNAME="ProductX"} (PROJECT)) $\rhd\lhd$ \textsubscript{(PNUMBER),(PNO)} (WORKS\_ON) \\
EMP\_WORK\_10 $\Leftarrow$ (EMPLOYEE) $\rhd\lhd$ \textsubscript{(SSN),(ESSN)} ( $\sigma$ \textsubscript{HOURS>10} (EMP\_W\_X)) \\
RESULT $\Leftarrow$ $\pi$ \textsubscript{LNAME,FNAME} ( $\sigma$ \textsubscript{DNO=5} (EMP\_WORK\_10))

\subsubsection*{Result}
\begin{center}
\begin{tabular}{ c | c }
  LNAME & FNAME \\ \hline
  Smith & John \\
  English & Joyce \\
\end{tabular}
\end{center}

\subsubsection*{(b) List the names of employees who have a dependent with the same first name as themselves.}
\subsubsection*{Relational Operators}
E $\Leftarrow$ (EMPLOYEE) $\rhd\lhd$ \textsubscript{(SSN,FNAME),(ESSN,DEPENDENT\_NAME)} (DEPENDENT)\\
R $\Leftarrow$ $\pi$ \textsubscript{LNAME,FNAME} (E)

\subsubsection*{Result}
\begin{center}
\begin{tabular}{ c | c }
  LNAME & FNAME \\ \hline
\end{tabular}
\end{center}

\subsubsection*{(c) Find the names of employees that are directly supervised by "Franklin Wong".}
\subsubsection*{Relational Operators}
WONG\_SSN $\Leftarrow$ $\pi$ \textsubscript{SSN} ( $\sigma$ \textsubscript{FNAME="Franklin" AND LNAME="Wong"} (EMPLOYEE))\\
WONG\_EMPS $\Leftarrow$ (EMPLOYEE) $\rhd\lhd$ \textsubscript{(SUPERSSN),(SSN)} (WONG\_SSN)\\
RESULT $\Leftarrow$ $\pi$ \textsubscript{LNAME,FNAME} (WONG\_EMPS)

\subsubsection*{Result}
\begin{center}
\begin{tabular}{ c | c }
  LNAME & FNAME \\ \hline
  Smith & John \\
  Narayan & Ramesh \\
  English & Joyce \\
\end{tabular}
\end{center}

\subsubsection*{(d) For each project, list the project name and the total hours per week (by all employees) spent on that project.}
\subsubsection*{Relational Operators}
PROJ\_HOURS(PNO,TOT\_HRS) $\Leftarrow$ \textsubscript{PNO} $\mathfrak{F}$ \textsubscript{SUM HOURS} (WORKS\_ON) \\
RESULT $\Leftarrow$ $\pi$ \textsubscript{PNAME,TOT\_HRS} ( (PROJ\_HOURS) $\rhd\lhd$ \textsubscript{(PNO),(PNUMBER)} (PROJECT) )

\subsubsection*{Result}  
\begin{center}
\begin{tabular}{ c | c }
  PNAME & TOT\_HRS  \\ \hline
  ProductX & 52.5 \\
  ProductY & 37.5 \\
  ProductZ & 50.0 \\
  Computerization & 55.0 \\
  Reorganization & 25.0 \\
  Newbenefits & 55.0 \\
\end{tabular}
\end{center}

\subsubsection*{(e) Retrieve the names of employees who work on every project.}
\subsubsection*{Relational Operators}
PROJ\_EMPS(PNO,SSN) $\Leftarrow$ $\pi$ \textsubscript{PNO,ESSN} (WORKS\_ON) \\
ALL\_PROJS(PNO) $\Leftarrow$ $\pi$ \textsubscript{PNUMBER} (PROJECT) \\
EMPS\_ALL\_PROJS $\Leftarrow$ PROJ\_EMPS $\div$ ALLPROJS \\
RESULT $\Leftarrow$ $\pi$ \textsubscript{LNAME,FNAME} (EMPLOYEE * EMP\_ALL\_PROJS)

\subsubsection*{Result}
\begin{center}
\begin{tabular}{ c | c }
  LNAME & FNAME \\ \hline
\end{tabular}
\end{center}

\subsubsection*{(f) Retrieve the names of employees who do not work on any project.}
\subsubsection*{Relational Operators}
ALL\_EMPS $\Leftarrow$ $\pi$ \textsubscript{SSN} (EMPLOYEE) \\
WORKING\_EMPS(SSN) $\Leftarrow$ $\pi$ \textsubscript{ESSN} (WORKS\_ON)\\
NON\_WORKING\_EMPS $\Leftarrow$ ALL\_EMPS - WORKING\_EMPS\\
RESULT $\Leftarrow$ $\pi$ \textsubscript{LNAME,FNAME} (EMPLOYEE * NON\_WORKING\_EMPS)

\subsubsection*{Result}
\begin{center}
\begin{tabular}{ c | c }
  LNAME & FNAME \\ \hline
\end{tabular}
\end{center}

\subsubsection*{(g) For each department, retrieve the department name and the average salary of all employees working in that department.}
\subsubsection*{Relational Operators}
DEPT\_AVG\_SALS(DNUMBER,AVG\_SAL) $\Leftarrow$ \textsubscript{DNO} $\mathfrak{F}$ \textsubscript{AVG SALARY} (EMPLOYEE)\\
RESULT $\Leftarrow$ $\pi$ \textsubscript{DNUMBER,AVG\_SAL} ( DEPT\_AVG\_SALS * DEPARTMENT )

\subsubsection*{Result}
\begin{center}
\begin{tabular}{ c | c }
  DNUMBER & AVG\_SAL  \\ \hline
  Research & 33250 \\
  Administration & 31000 \\
  Headquarters & 55000 \\
\end{tabular}
\end{center}

\subsubsection*{(h) Retrieve the average salary of all female employees.}
\subsubsection*{Relational Operators}
RESULT(AVG\_F\_SAL) $\Leftarrow$ $\mathfrak{F}$ \textsubscript{AVG SALARY} ( $\sigma$ \textsubscript{SEX="F"} (EMPLOYEE) )

\subsubsection*{Result}
\begin{center}
\begin{tabular}{ c }
  AVG\_F\_SAL \\ \hline
  31,000 \\
\end{tabular}
\end{center}

\subsubsection*{(i) Find the names and addresses of employees who work on at least one project located in Houston but whose department has no location in Houston.}
\subsubsection*{Relational Operators}
E\_P\_HOU(SSN) $\Leftarrow$ $\pi$ \textsubscript{ESSN} (WORKS\_ON $\rhd\lhd$ \textsubscript{(PNO),(PNUMBER)} ( $\sigma$ \textsubscript{PLOCATION="Houston"} (PROJECT)))\\
D\_NO\_HOU $\Leftarrow$ $\pi$ \textsubscript{DNUMBER} (DEPARTMENT) - $\pi$ \textsubscript{DNUMBER} ( $\sigma$ \textsubscript{DLOCATION="Houston"} (DEPARTMENT))\\
E\_D\_NO\_HOU $\Leftarrow$ $\pi$ \textsubscript{SSN} (EMPLOYEE  $\rhd\lhd$ \textsubscript{(PNO),(DNUMBER)} (D\_NO\_HOU))\\
RESULT\_EMPS $\Leftarrow$ E\_P\_HOU - E\_D\_NO\_HOU \\
RESULT $\Leftarrow$ $\pi$ \textsubscript{LNAME,FNAME,ADDRESS} (EMPLOYEE * RESULT\_EMPS)

\subsubsection*{Result}
\begin{center}
\begin{tabular}{ c | c | c}
  LNAME & FNAME  & ADDRESS \\ \hline
  Wallace & Jennifer & 291 Berry, Bellaire, TX \\
\end{tabular}
\end{center}

\subsubsection*{(j) List the last names of department managers who have no dependents.}
\subsubsection*{Relational Operators}
DEPT\_MANAGERS(SSN)<-- $\pi$ \textsubscript{MGRSSN} (DEPARTMENT) \\
EMPS\_WITH\_DEPENDENTS(SSN) $\Leftarrow$ $\pi$ \textsubscript{ESSN} (DEPENDENT) \\
RESULT\_EMPS $\Leftarrow$ DEPT\_MANAGERS - EMPS\_WITH\_DEPENDENTS \\
RESULT $\Leftarrow$ $\pi$ \textsubscript{LNAME,FNAME} (EMPLOYEE * RESULT\_EMPS)

\subsubsection*{Relational Operators}
\begin{center}
\begin{tabular}{ c | c }
  LNAME & FNAME \\ \hline
  Borg & James \\
\end{tabular}
\end{center}