\section*{Chapter 3}
\subsection*{3.11 Suppose that each of the following Update operations is applied directly to the database state shown in Figure 3.6. Discuss all integrity constraints violated by each operation, if any, and the different ways of enforcing these constraints.}
\subsubsection*{(a) Insert <"Robert", "F", "Scott", "943775543", "1972-06-21", "2365 Newcastle Rd, Bellaire, TX", M, 58000, "888665555", 1> into EMPLOYEE.}
No constraints violated.

\subsubsection*{(b) Insert <"ProductA", 4, "Bellaire", 2> into PROJECT.}
This insertion would Violates referential integrity because DNUM=2 and there is no tuple in the DEPARTMENT relation with DNUMBER=2. We may enforce the constraint by: (i) rejecting the insertion of the new PROJECT tuple, (ii) changing the value of DNUM in the new PROJECT tuple to an existing DNUMBER value in the DEPARTMENT relation, or (iii) inserting a new DEPARTMENT tuple with DNUMBER=2.

\subsubsection*{(c) Insert <"Production", 4, "943775543", "2007-10-01"> into DEPARTMENT.}
Violates both the key constraint and referential integrity. Violates the key constraint because there already exists a DEPARTMENT tuple with DNUMBER=4. We may enforce this constraint by: (i) rejecting the insertion, or (ii) changing the value of DNUMBER
in the new DEPARTMENT tuple to a value that does not violate the key constraint. Violates referential integrity because MGRSSN="943775543" and there is no tuple in the EMPLOYEE relation with SSN="943775543". We may enforce the constraint by: (i) rejecting the insertion, (ii) changing the value of MGRSSN to an existing SSN value in EMPLOYEE, or (iii) inserting a new EMPLOYEE tuple with SSN="943775543".

\subsubsection*{(d) Insert <"677678989", NULL, "40.0"> into WORKS\_ON.}
(d) Violates both the entity integrity and referential integrity. Violates entity integrity because PNO, which is part of the primary key of WORKS\_ON, is null. We may enforce this constraint by: (i) rejecting the insertion, or (ii) changing the value of PNO in the new WORKS\_ON tuple to a value of PNUMBER that exists in the PROJECT relation. Violates referential integrity because ESSN="677678989" and there is no tuple in the EMPLOYEE relation with SSN="677678989". We may enforce the constraint by: (i) rejecting the insertion, (ii) changing the value of ESSN to an existing SSN value in EMPLOYEE, or (iii) inserting a new EMPLOYEE tuple with SSN="677678989".

\subsubsection*{(e) Insert <"453453453","John","M","1990-12-12","spouse"> into DEPENDENT.}
Nope! Not a single constraint was violated.

\subsubsection*{(f) Delete the WORKS\_ON tuples with Essn = "333445555".}
No constraints violated in the making of this delete.

\subsubsection*{(g) Delete the EMPLOYEE tuple with Ssn = "987654321".}
(g) Violates referential integrity because several tuples exist in the WORKS\_ON, DEPENDENT, DEPARTMENT, and EMPLOYEE relations that reference the tuple being deleted from EMPLOYEE. We may enforce the constraint by: (i) rejecting the deletion, or (ii) deleting all tuples in the WORKS\_ON, DEPENDENT, DEPARTMENT, and EMPLOYEE relations whose values for ESSN, ESSN, MGRSSN, and SUPERSSN, respectively, is equal to "987654321".

\subsubsection*{(h) Delete the PROJECT tuple with Pname="ProductX".}
(h) Violates referential integrity because two tuples exist in the WORKS\_ON relations that reference the tuple being deleted from PROJECT. We may enforce the constraint by: (i) rejecting the deletion, or (ii) deleting the tuples in the WORKS\_ON relation whose value for PNO=1 (the value for the primary key PNUMBER for the tuple being deleted from PROJECT).

\subsubsection*{(i) Modify the Mgr\_ssn and Mgr\_start\_date of the DEPARTMENT tuple with Dnumber = 5 to "123456789" and "2007-10-01", respectively.}
No constraints violated.

\subsubsection*{(j) Modify the Super\_ssn attribute of the EMPLOYEE tuple with Ssn = "999887777" to "943775543".}
Violates referential integrity because the new value of SUPERSSN="943775543" and there is no tuple in the EMPLOYEE relation with SSN="943775543". We may enforce the constraint by: (i) rejecting the deletion, or (ii) inserting a new EMPLOYEE tuple with SSN="943775543".

\subsubsection*{(k) Modify the Hours attribute of the WORKS\_ON tuple with Essn = "999887777" and Pno = 10 to "5.0".}
No constraints violated.